Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-11-28 presentations/presentation-illustrations.tex

\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{positioning}
\usetikzlibrary{patterns}
\usetikzlibrary{decorations.pathreplacing}
\usepackage{tikzpeople}

@@

@@

\begin{frame}
  \frametitle{Stack and return capabilities: another attack}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
      \draw[draw=gray!80] (0,0) -- (0,15);
      \draw[draw=gray!80] (6,0) -- (6,15);
      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
    \end{scope}

@@

    % attacker calls trusted code again
    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.>[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.>[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw<.> node[right=0cm of sp1] { stack pointer };
    \draw<.>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.> (6,13.25) edge[red,thick,->,in=45] (spold);

    % attacker calls trusted code again
    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.>[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.>[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \draw<.>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
    \draw<.> (3,10.5) node[devil,mirrored,minimum size=1cm] {};
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw<.> node[right=0cm of sp1] { stack pointer };
    \draw<.>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.> (6,13.25) edge[red,thick,->,in=45] (spold);
    \draw<.> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Stack Capabilities Problems}
  \begin{itemize}
  \item Stack capabilities not a panacea
  \end{itemize}
\end{frame}
\end{document}

>>>>

\usepackage{pgf}
\usepackage{ulem}
\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{positioning}
\usetikzlibrary{patterns}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{arrows,shapes}
\usepackage{tikzpeople}

@@

  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Stack and return capabilities: Attack 1}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
      \draw[draw=gray!80] (0,0) -- (0,15);
      \draw[draw=gray!80] (6,0) -- (6,15);
      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
    \end{scope}

@@

\begin{frame}
  \frametitle{Stack and return capabilities: Attack 2}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
      \draw[draw=gray!80] (0,0) -- (0,15);
      \draw[draw=gray!80] (6,0) -- (6,15);
      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
    \end{scope}

@@

    % attacker calls trusted code again
    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.>[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.>[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15);
      \draw<.>[fill=white] (0,9) rectangle (6,15.1);
    \end{scope}
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw<.> node[right=0cm of sp1] { stack pointer };
    \draw<.>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.> (6,13.25) edge[red,thick,->,in=45] (spold);

    % attacker calls trusted code again
    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.>[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.>[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \draw<.>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15);
      \draw<.>[fill=white] (0,12) rectangle (6,15.1);
    \end{scope}
    \draw<.> (3,10.5) node[devil,mirrored,minimum size=1cm] {};
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw<.> node[right=0cm of sp1] { stack pointer };
    \draw<.>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.> (6,13.25) edge[red,thick,->,in=45] (spold);
    \draw<.> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Stack Capabilities Problems}
  \begin{itemize}
  \item Stack capabilities are not a panacea
  \item Stack pointers must not leave stack frame (through memory or stack)
  \item Similar problems for return pointers (omitted)
  \item Solution: local capabilities?
    \begin{itemize}
    \item Make stack and return capabilities local
    \item Cannot leave registers (except on stack) (\sout{Attack 1})
    \item Stack clearing on boundary crossings (\sout{Attack 2})
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Local Stack Capabilities prevent Attack 1}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
      \draw[draw=gray!80] (0,0) -- (0,15);
      \draw[draw=gray!80] (6,0) -- (6,15);
      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
    \end{scope}

    \draw<.>[green,thick] (13,7.5) -- (19,9);
    \draw<.>[green,very thick] (19,7.5) -- (13,9);
    \draw node[green,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw] () at (14,4) {Stack pointer is local!};

  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Stack Clearing prevent Attack 2}
  TODO
\end{frame}
\end{document}
